---
title: "eds221day7interactive"
author: "carly"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(palmerpenguins)
```

## *Data Wrangling Refresher*

Starting from penguins...

1.  *Only include penguins at dream and biscoe islands*

    ```{r}
    penguins %>% 
      filter(island == "Dream" | island == "Biscoe")
    ```

*2. Remove the year and sex variables*

```{r}
penguins_removed <- penguins %>%
  select(-year, -sex)
```

3.  *Add a new column body_mass_kg with penguin mass converted from g to kg*

    ```{r}
    body_mass_kg <- penguins %>% 
      mutate(body_mass_kg = body_mass_g / 1000)
    ```

4.  *Rename the island variable to location*

    ```{r}
    penguins %>% 
      rename("location" = "island")
    ```

Starting from penguins, write a single piped sequence to:

-limit to adelie penguins

-remove obs where flipper length mm is NA (drop_na())

-group the data by sex

-create a summary table that contains the mean, sd, and sample size of flipper length

```{r}
penguins %>% 
  filter(species == "Adelie") %>% 
  drop_na(flipper_length_mm, sex) %>% 
  group_by(sex) %>% 
  summarise(mean_flipper = mean(flipper_length_mm), sd_flipper = sd(flipper_length_mm), count = n())
```

Another way to get the sample size of groups is to use count...

```{r}
penguins %>% count(species, island, sex, name = "count")
```
